---
title: "Analysis of modifiers of the word machine for project Living Machines 2"
author: "Barbara McGillivray"
date: "28 July 2021"
output: word_document
---


# Initialization

## Load libraries and data, and set initial parameters

Load libraries

```{r message=FALSE}
op <- options(warn = (-1)) # suppress warnings 
library(gplots)
library(ggplot2)
library(gdata)
library(stopwords)
library(tm)
options(op) # reset the default value
```


Directory names (to be customised):
  
```{r Initial parameters}
path_in =  file.path("/Users", "barbaramcgillivray", "OneDrive - The Alan Turing Institute", "Research", "2021", "LwM", "Living machines2", "Data", fsep = "/")
path_data = paste(path_in, "older/", sep = "/")
path_plots = paste(path_in, "plots", sep = "/")
path_out = file.path("/Users", "barbaramcgillivray", "OneDrive - The Alan Turing Institute", "Research", "2021", "LwM", "Living machines2", "Machine attributes analysis")
```

Corpus names:
  
```{r message=FALSE, warning = FALSE}
corpora = c("blb", "hmd", "jsa", "rsc")
```

# Read files

NB: This takes a long time. 

```{r message=FALSE, warning = FALSE, eval=FALSE}
datasets <- list()
count = 0
for (corpus in corpora) {
  print(corpus)
  count = count + 1
  file_name = paste(corpus, "_withmere.tsv", sep = "")
  dataset = read.csv(paste(path_data, file_name, sep = "/"), header = T, sep = "\t", colClasses=c("synt"="character"))
  print(dim(dataset))
  # only keep relevant columns:
  if (corpus == "blb"){
    cols = c("identifier", "date", "currentSentence", "synt")
  }
  else if (corpus == "hmd"){
    cols = c("item_code", "year", "currentSentence", "synt")
  }
  else if (corpus == "jsa"){
    cols= c("filename", "year", "currentSentence", "synt")
  }
  else {
    cols= c("volume", "year", "currentSentence", "synt")
  }
  dataset = dataset[,cols]
  if (corpus == "blb"){
    colnames(dataset)[2] <- "year"
  }
  # add to list of datasets:
  datasets[[count]] = dataset
}
```

Overview of the datasets:

```{r}
for (dataset in datasets){
  print(names(dataset))
}
```

Summaries:

```{r}
for (dataset in datasets){
print(summary(dataset$year))
  hist(dataset$year)
}
```

## Extract modifiers of "machine"


See machine_attributes_analysis.R.

# BL Books corpus

```{r}
machine_dep_blb = read.csv(paste(path_out, 'machine_dep_blb.csv', sep = "/"))
dim(machine_dep_blb)
```

Summary:

```{r}
summary(machine_dep_blb)
```

Distribution over the years:

```{r}
hist(machine_dep_blb$year, title = "Distribution of years in dataset")
```


Exclude punctuation marks, symbols, determiners, adpositions, conjunctions and unexpected PoS values: 

```{r}
machine_dep_blb1 = machine_dep_blb[machine_dep_blb$pos != "" & machine_dep_blb$pos != "_" & machine_dep_blb$pos !="-" & machine_dep_blb$pos != "PUNCT" & machine_dep_blb$pos != "SYM" & machine_dep_blb$pos != "DET" & machine_dep_blb$pos != "ADP" & machine_dep_blb$pos != "CCONJ"  & machine_dep_blb$pos != "SCONJ"  & machine_dep_blb$pos != "X" & machine_dep_blb$pos != "PART" & machine_dep_blb$pos != "INTJ"  & machine_dep_blb$pos != "AUX" & !grepl("^[[:digit:]]",machine_dep_blb$pos) & machine_dep_blb$pos !='\"meansa' & machine_dep_blb$pos != "^card",]
head(machine_dep_blb1)
```

Part of speech of dependants of "machine":

```{r}
machine_dep_blb_pos.df = as.data.frame(table(machine_dep_blb1$pos))
colnames(machine_dep_blb_pos.df) = c("Pos", "Freq")
machine_dep_blb_pos.df = machine_dep_blb_pos.df[order(-machine_dep_blb_pos.df$Freq),]
machine_dep_blb_pos.df = machine_dep_blb_pos.df[machine_dep_blb_pos.df$Freq > 0,]
head(machine_dep_blb_pos.df)
```

Barplot:

```{r}
barplot(machine_dep_blb_pos.df$Freq, names.arg = machine_dep_blb_pos.df$Pos, las = 2)
```

Combine dependants and PoS info:

```{r}
machine_dep_blb.df = as.data.frame(table(machine_dep_blb1$pos, machine_dep_blb1$dependant))
```

Exclude punctuation marks, symbols, determiners, adpositions, conjunctions and unexpected PoS values: 

```{r}
machine_dep_blb_pos_dep = machine_dep_blb.df[machine_dep_blb.df$Freq>0 & machine_dep_blb.df$Var1 != "" & machine_dep_blb.df$Var2 != "" & machine_dep_blb.df$Var1 != "_" & machine_dep_blb.df$Var2 != "_" & machine_dep_blb.df$Var2!="-" & machine_dep_blb.df$Var1 != "PUNCT" & machine_dep_blb.df$Var1 != "SYM" & machine_dep_blb.df$Var1 != "DET" & machine_dep_blb.df$Var1 != "ADP" & machine_dep_blb.df$Var1 != "CCONJ"  & machine_dep_blb.df$Var1 != "SCONJ"  & machine_dep_blb.df$Var1 != "X" & machine_dep_blb.df$Var1 != "PART" & machine_dep_blb.df$Var1 != "INTJ"  & machine_dep_blb.df$Var1 != "AUX" & !grepl("^[[:digit:]]",machine_dep_blb.df$Var1) & machine_dep_blb.df$Var1 !='\"meansa' & machine_dep_blb.df$Var1 != "^card",]
colnames(machine_dep_blb_pos_dep) = c("PoS", "Dependant", "Freq")
head(machine_dep_blb_pos_dep)
```

Sort by PoS and decreasing frequency:

```{r}
machine_dep_blb_pos_dep = machine_dep_blb_pos_dep[order(machine_dep_blb_pos_dep$PoS, -machine_dep_blb_pos_dep$Freq),]
head(machine_dep_blb_pos_dep)
```

Save to file:

```{r}
write.csv(machine_dep_blb_pos_dep, paste(path_out, paste('machine_dep_blb_pos_dep.csv', sep = ""), sep = "/"), row.names = F)
```

For each part of speech, I show the top 50 words appearing as dependant of machine:

```{r}
top_per_pos = vector(mode = "list", length = nrow(unique(machine_dep_blb_pos_dep[c("PoS")])))
i = 0
for (pos in unique(machine_dep_blb_pos_dep[[c("PoS")]])){
  print(pos)
  i = i + 1
  this_pos = machine_dep_blb_pos_dep[machine_dep_blb_pos_dep$PoS==pos,c("Dependant")]
  this_pos = droplevels(this_pos)
  top_per_pos[[i]] = head(this_pos,50)
}
top_per_pos.df = data.frame(top_per_pos)
colnames(top_per_pos.df) = unique(machine_dep_blb_pos_dep[[c("PoS")]])
top_per_pos.df
```


Save to file:

```{r}
write.csv(top_per_pos.df, paste(path_out, paste('top_dependants_per_pos_blb.csv', sep = ""), sep = "/"), row.names = F)
```

# Other corpora

```{r}
machine_dep = read.csv(paste(path_out, 'machine_dep.csv', sep = "/"))
dim(machine_dep)
```

Summary:

```{r}
summary(machine_dep)
```


Distribution over the corpora:

```{r}
table(machine_dep$corpus)
```


Distribution over the years:

```{r}
hist(machine_dep$year, title = "Distribution of years in dataset")
ggplot(machine_dep, aes(x = year, fill = corpus)) + 
  geom_histogram()
```

I define a function that does the analysis above for each corpus:

```{r}
extract_machine_dependants <- function(corpus) {
  print("Loading data...")
  machine_dep_c = machine_dep[machine_dep$corpus==corpus,]

  #Exclude punctuation marks, symbols, determiners, adpositions, conjunctions and unexpected PoS values: 
  print("Clean pos...")
  machine_dep_c1 = machine_dep_c[machine_dep_c$dependant != "" & machine_dep_c$pos != "" & machine_dep_c$pos != "_" & machine_dep_c$pos !="-" & machine_dep_c$pos != "PUNCT" & machine_dep_c$pos != "SYM" & machine_dep_c$pos != "DET" & machine_dep_c$pos != "ADP" & machine_dep_c$pos != "CCONJ"  & machine_dep_c$pos != "SCONJ"  & machine_dep_c$pos != "X" & machine_dep_c$pos != "PART" & machine_dep_c$pos != "INTJ"  & machine_dep_c$pos != "AUX" & !grepl("^[[:digit:]]",machine_dep_c$pos) &  grepl("^[A-Z]+$",machine_dep_c$pos),]
  
  #Part of speech of dependants of "machine":
  print("Table with PoS...")
machine_dep_c_pos.df = as.data.frame(table(machine_dep_c1$pos))
colnames(machine_dep_c_pos.df) = c("Pos", "Freq")
machine_dep_c_pos.df = machine_dep_c_pos.df[order(-machine_dep_c_pos.df$Freq),]
machine_dep_c_pos.df = machine_dep_c_pos.df[machine_dep_c_pos.df$Freq > 0,]

# barplot:
print("Barplot...")
print(paste(path_out,paste("barplot_machine_dep_c_pos_", corpus, ".png", sep = ""), sep = "/"))
png(paste(path_out,paste("barplot_machine_dep_c_pos_", corpus, ".png", sep = ""), sep = "/")) 
barplot(machine_dep_c_pos.df$Freq, names.arg = machine_dep_c_pos.df$Pos, las = 2)
dev.off()

#Combine dependants and PoS info:
print("Table with dependants and pos...")
machine_dep_c.df = as.data.frame(table(machine_dep_c1$pos, machine_dep_c1$dependant))
#Exclude punctuation marks, symbols, determiners, adpositions, conjunctions and unexpected PoS values: 
machine_dep_c_pos_dep = machine_dep_c.df[machine_dep_c.df$Freq>0 & machine_dep_c.df$Var1 != "" & machine_dep_c.df$Var2 != "" & machine_dep_c.df$Var1 != "_" & machine_dep_c.df$Var2 != "_" & machine_dep_c.df$Var2!="-" & machine_dep_c.df$Var1 != "PUNCT" & machine_dep_c.df$Var1 != "SYM" & machine_dep_c.df$Var1 != "DET" & machine_dep_c.df$Var1 != "ADP" & machine_dep_c.df$Var1 != "CCONJ"  & machine_dep_c.df$Var1 != "SCONJ"  & machine_dep_c.df$Var1 != "X" & machine_dep_c.df$Var1 != "PART" & machine_dep_c.df$Var1 != "INTJ"  & machine_dep_c.df$Var1 != "AUX" & !grepl("^.*[[:digit:]]",machine_dep_c.df$Var1) & machine_dep_c.df$Var1 !='\"meansa' & machine_dep_c.df$Var1 != "^card" & !grepl("^[-;:.].*$",machine_dep_c.df$Var1) & !grepl('^.*\"$',machine_dep_c.df$Var1) & !grepl("^[—].*$",machine_dep_c.df$Var1) ,]
colnames(machine_dep_c_pos_dep) = c("PoS", "Dependant", "Freq")

#Sort by PoS and decreasing frequency:
print("Sorting table...")
machine_dep_c_pos_dep = machine_dep_c_pos_dep[order(machine_dep_c_pos_dep$PoS, -machine_dep_c_pos_dep$Freq),]

#Save to file:
print("Saving file...")
write.csv(machine_dep_c_pos_dep, paste(path_out, paste('machine_dep_', corpus, '_pos_dep.csv', sep = ""), sep = "/"), row.names = F)

#For each part of speech, I show the top 50 words appearing as dependant of machine:
print("Top per pos...")
top_per_pos = vector(mode = "list", length = nrow(unique(machine_dep_c_pos_dep[c("PoS")])))
i = 0
for (pos in unique(machine_dep_c_pos_dep[[c("PoS")]])){
  print(pos)
  i = i + 1
  this_pos = machine_dep_c_pos_dep[machine_dep_c_pos_dep$PoS==pos,c("Dependant")]
  this_pos = droplevels(this_pos)
  top_per_pos_i = rep(NA, 50)
  #top_per_pos[[i]] = head(this_pos,50)
  top_per_pos_i[1:length(head(this_pos,50))] = head(as.character(this_pos),50)
  top_per_pos[[i]] = top_per_pos_i
}
top_per_pos.df = data.frame(top_per_pos)
colnames(top_per_pos.df) = unique(machine_dep_c_pos_dep[[c("PoS")]])
top_per_pos.df
#Save to file:
print("Saving file")
write.csv(top_per_pos.df, paste(path_out, paste('top_dependants_per_pos_', corpus, '.csv', sep = ""), sep = "/"), row.names = F)
}
```

Run the function on all corpora:

```{r}
for (c in corpora){
  print(c)
extract_machine_dependants(c)
}
```

Now I compare the adjective modifiers of machines across the corpora:

```{r}
top_per_pos_all_adj = matrix(, nrow = 50, ncol = length(corpora))
for (i in 1:length(corpora)){
c = corpora[i]
top_per_pos = read.csv(paste(path_out,  paste('top_dependants_per_pos_', c, '.csv', sep = ""), sep = "/"))
top_per_pos_adj = as.character(top_per_pos[,c("ADJ")])
top_per_pos_all_adj[,i] = top_per_pos_adj
}
top_per_pos_all_adj.df = data.frame(top_per_pos_all_adj)
names(top_per_pos_all_adj.df) = corpora
top_per_pos_all_adj.df
print("Saving file")
write.csv(top_per_pos_all_adj.df, paste(path_out, paste('top_per_pos_all_adj.df.csv', sep = ""), sep = "/"), row.names = F)
```



I create a table with the presence of words in each corpus:

```{r}
data.frame(table(top_per_pos_all_adj.df))
```


